%
% Cosmic Raysssss (chap1.tex)
%
\chapter{Cosmic Ray Astrophysics}\label{chap:cosmicrays}
Cosmic rays are ultra-relativistic charged particles of astrophysical origin.
They were discovered in 1912 when Victor Hess, armed with his electroscope and an appetite for scientific understanding, performed several hot-air balloon flights to study radiation in the atmosphere.
Hess measured increasing levels of atmospheric radiation as his balloon increased in altitude, leading him to hypothesize the existence of a ``radiation of cosmic origin'', to which he gave the name cosmic rays.
He was later awarded the physics Nobel prize in 1936 for his discovery and the misnomer given to these particles stuck.

Cosmic rays have been studied for over a century since their discovery, yet the acceleration mechanisms producing ultra-high-energy cosmic rays, those above \qty{1}{\exa\electronvolt}\,=\,\qty[parse-numbers=false]{10^{18.0}}{\electronvolt} in energy, and their astrophysical sources are still a mystery.
There are several measurable properties of cosmic rays that help to provide progress towards answering this mystery including the energy, mass, and arrival direction at Earth of the cosmic ray.
Furthermore, cosmic rays are the highest energy subatomic particles in the known Universe, with energies exceeding the most energetic collisions at the Large Hadron Collider (LHC) by nearly three orders of magnitude.
Hence, ultra-high-energy cosmic rays push the frontiers of both astrophysics and particle physics simultaneously.
This chapter will mostly discuss the astrophysical implications of cosmic-ray physics and~\Cref{chap:airshowers} will present the particle physics components; however, they are inherently coupled, with aspects of each physics sub-field in both chapters.


%terrestrial particle accelerators by nearly three orders of magnitude.

%with center of mass energies of collisions with stationary atmospheric particles exceeding the \qty{14}{\tera\electronvolt} center of mass energy of the proton-proton collisions at the Large Hadron Collider (LHC)
%What makes cosmic rays interesting beyond their potential for astrophysical source and process understanding is 
%Not only do cosmic rays probe the 

%%%%%%% Maybe use the above comments???

%Cosmic rays were discovered in 1912 when Victor Hess, armed with his electroscope and an appetite for scientific understanding, performed several hot-air balloon flights to study atmospheric ionizing radiation.
%As Hess' balloon increased in altitude, he measured higher levels of atmospheric radiation, leading him to hypothesize the existence of a ``radiation of cosmic origin'' which he called cosmic rays.
%This ``cosmic radiation'' was the electroscope measuring the increasing flux of astrophysical, ultra-relativistic, charged particles with altitude.
%Hess did not realize it at the time, but his electroscope was measuring the increasing flux of ultra-relativistic charged particles with altitude.

%Hess was later awarded the physics Nobel prize in 1936 for his discovery.
%Physicists continued studying cosmic rays over the last century determining they are, contrary to what their misnomer suggests, charged particles of astrophysical origin.

%Like light, cosmic rays carry information throughout the Universe from where they came.
%However, 

%Even though 
%Hess was awarded the Nobel prize in physics in 1936 for his discovery.

%The Universe is beaming with astrophysical processes occurring at various energy scales.
%At each energy scale, the physical processes occurring at these astrophysical objects produce charged particles, referred to as cosmic rays.
%Yet, 




%Cosmic rays span over ten orders of magnitude in energy, with the highest energy cosmic rays exceeding the center of mass energy of the Large Hadron Collider (LHC) collisions by several orders of magnitude.


%However, after more than 100 years of cosmic-ray research, th 

%this ionizing radiation is of cosmic origin.
%He named this type of radiation ``cosmic rays'' and was awarded the physics Nobel prize in 1936 for his discovery.
%After over 100 years of studying cosmic rays, physicists now understand  
%which he named ``cosmic rays'' 

%During his flights, he measured higher levels of radiation as he went higher in the atmosphere.

%Over 100 years ago, Victor Hess 

%Today, we know that cosmic rays are 

%inadvertently discovered 
%For centuries, physicists and the like have studied radiation, charges, and similar phenomena.
%However, radiation was always assumed to be from the ground.
%Until XXXX when SOME GUY proposed the idea that radiation could come from space.

%The first instrument used to study these processes was the electroscope, a simple device used to detect the presence of charge.

%designed to detect the presence of charge was the electroscope.

%Studies were conducted using electroscopes, the first instrument designed to detect the presence of charge, to prove various properties of radiation and radiactive materials.
%Marie and Pierre Curie 
%Yet, over 100 years ago, some physicsits had the idea that 

%The Universe is beaming with astrophyscal processes occurring at various energy scales.
%At each energy scale, the physical processes occurring at these astrophysical objects produce charged particles, referred to as cosmic rays.
%Yet, 

%Cosmic rays span over ten orders of magnitude in energy, with the highest energy cosmic rays exceeding the center of mass energy of the Large Hadron Collider (LHC) collisions by several orders of magnitude.



\section{Energy Spectrum}\label{sec:spectrum}
Cosmic rays follow a steeply declining energy spectrum.

\section{Mass}\label{sec:mass}
Directly related to determining the cosmic-ray energy is also determining its mass.

\section{Anisotropy}\label{sec:anisotropy}
Cosmic rays serve as a probe to the deepest depths of the known Universe.

\section{Sources, Acceleration Mechanisms, and MM}\label{sec:sources}
In addition to cosmic rays, their acceleration sites also produce other particle without charge and high-energy electromagnetic radiation in the form of gamma rays.

\section{Propagation Effects}\label{sec:propagation}
Further complicating the source determination of cosmic rays, and contributing to the generation of neutral particles along the path of travel to Earth, are propagation effects.