%
% The Pierre Auger Observatory (chapter 3)
%
\chapter{The Pierre Auger Observatory}\label{chap:auger}

The Pierre Auger Observatory is located near the town of Malarg\"ue in the Mendoza province of Argentina between the Andes mountain range and la Pampa Amarilla.
Auger, as the observatory is commonly referred to, is the world's largest cosmic-ray observatory, covering a total area of \qty{3000}{\kilo\meter^{2}}.
In context, this is about the size of the state of Rhode Island, or for those with a more worldly perspective, slightly larger than the size of Luxembourg.
The observatory's size allows the detection of ultra-high-energy cosmic rays, starting at energies of $\approx$\qty[parse-numbers=false]{10^{17}}{\electronvolt}, with high statistics.
However, the observatory is ground based and therefore indirectly detects the highest energy particles in the known Universe from the air showers they initiate after colliding with atmospheric nuclei.

Not only is Auger the world's largest cosmic-ray observatory, but adding to the novelty is the hybrid detection mode employed by the observatory.
The hybrid design allows simultaneous air-shower observations with two separate, yet complimentary techniques.
The observatory consists of two primary detectors, the surface detector (SD), an array of over 1600 water-Cherenkov detectors (WCDs) spread across the full \qty{3000}{\kilo\meter^{2}} footprint of the observatory, and a fluorescence detector (FD), 27 fluorescence telescopes overlooking the SD array at four locations around the edges of the array.
Auger has been recording data in hybrid mode for over 20 years, officially starting in 2004 and with full completion of the SD in 2008.
In March 2026, Auger finished deploying detectors for the next-generation upgrade of the observatory, AugerPrime.
An international funding agreement will keep the observatory in operation until at least 2035, allowing for high-quality air-shower measurements with AugerPrime, and the potential for prototyping detectors for future experiments in the Argentinian Pampa.
This chapter will describe the Pierre Auger Observatory and its detectors systems in detail.


\section{Surface Detector}\label{sec:sd}

The Pierre Auger Observatory SD consists of 1660 WCDs deployed on a triangular grid to cover the full surface area of the array.
The spacing of the grid is \qty{1500}{\meter}, designed such that all cosmic rays landing within the array grid and above an energy of \qty[parse-numbers=false]{10^{18.5}}{\electronvolt} are detected.
Within the SD-1500 are two nested triangular sub-arrays of WCDs with denser spacing to lower the energy threshold of cosmic-ray detection.
They are the SD-750, with \qty{750}{\meter} spacing between each detector and the SD-433, nested within the SD-750 and with a detector spacing of \qty{433}{\meter}.
The SD-750 consists of 61 WCDs covering an area of \qty{23.5}{\kilo\meter^{2}}, lowering the energy detection threshold to \qty[parse-numbers=false]{10^{17.5}}{\electronvolt}, while the SD-433 consists of 19 WCDs over an area of \qty{1.95}{\kilo\meter^{2}} and can detect air showers as low as \qty[parse-numbers=false]{10^{16.5}}{\electronvolt} in energy.
A map of the Pierre Auger Observatory, including all three SD arrays, is shown in INSERTFIGURE.

Each WCD is a cylinder of \qty{1.2}{\meter} height and \qty{3.6}{\meter} diameter filled with \qty{11000}{\liter} of ultra-pure water.
The water inside the WCD is contained in a bladder composed of Tyvek with a fully reflective inner surface.
At the top of each WCD are three, equally spaced, 9-inch phototmultiplier tubes (PMTs) to detect the Cherenkov radiation generated from ultra-relativistic secondary particles from cosmic-ray air showers traversing the ultra-pure water.
The PMT signals are digitized by a \qty{40}{\mega\hertz} flash analog-to-digital converter (FADC), corresponding to a bin width of \qty{25}{\nano\second} for the FADC conversion.
Two signals are digitized per PMT, the low-gain channel, taken directly from the PMT anode, and the high-gain channel, taken from the last PMT dynode with a nominal amplification of 32 times the anode channel, providing a total dynamic range of approximately 17 bits.
All processes are handled by the Auger Unified Board (UB), the main electronics board of each SD station and is described in more detail at REFERENCE.
Other components of the individual WCDs include a standalone communications system, for communications and data transfer to the Central Data Acquisition System (CDAS) of the observatory, and a solar power system for powering the station.
The solar power system allows an approximate 100\% uptime of the SD.

Signals from the PMTs are digitized in electronics units, integrated ADC units where the integration is performed over the bin width of the FADC.
The digitized signals are converted to units of Vertical Equivalent Muons (VEM) to provide a common reference unit across the array, where one VEM is defined as the total signal deposited in the WCD by a single muon traversing the WCD vertically and centrally.
The conversion to VEM is performed ``online'' (within the electronics of the station) at each WCD using the calibration histograms of the individual PMTs.
A WCD equipped with an RDC hodoscope was used to experimentally confirm the conversion (REF 6 Calibration).
The calibration histograms are generated from \qty{60}{\second} of atmospheric muon data recorded with the \texttt{muon\_buffer} of the local station.
See~\cref{app:layered_wcd} for further discussion of the \texttt{muon\_buffer}.

\subsection{SD Triggers}\label{sec:sd_triggers}

The atmospheric muons recorded for local station calibration arrive at an approximate rate of \qty{3}{\kilo\hertz}.
A tiered SD trigger system reduces this to a feasible rate for physics analyses.
To understand the threshold levels defined for the first two tiers of the trigger system an understanding of the peak and charge calibration histograms is needed.
The charge histogram is generated by integrating the total signals of individual atmospheric muons triggering the \texttt{muon\_buffer}, with charges ($Q$) measured in units of time-integrated FADC bins, or simply FADC bins.
The second maximum in the charge histogram is known as the ``muon peak'', where the charge value of this maximum, represented as $Q_{\rm VEM}^{\rm peak}$, defines the conversion constant from electronics units to VEM units.
Related to the charge histogram is the peak histogram, generated from the maximum amplitudes of the signal recorded by the \texttt{muon\_buffer}.
The peak histogram has units of FADC bins per \qty{25}{\nano\second} and is sometimes referred to as the current ($I$) histogram because the second maximum of this histogram, represented as $I_{\rm VEM}^{\rm peak}$, defines the end-to-end gain of the PMTs.
The end-to-end PMT gains are set individually and a local station-level calibration is done to continually adjust this value (see AUGER CALIBRATION SD for more information).

The first-level (T1) and second-level (T2) triggers are generated locally by the station electronics and depend directly on $I_{\rm VEM}^{\rm peak}$.
Two types of T1 trigger are possible, each designed for sensitivity to different particle types.
The first is a simple threshold trigger, Th-T1, requiring a coincidence of all three PMTs to measure a signal above 1.75$\,I_{\rm VEM}^{\rm peak}$.
This reduces the trigger rate of the SD from $\approx$\qty{3}{\kilo\hertz} to $\approx$\qty{100}{\hertz} and is more sensitive to muons, as these particles generate large amplitude and temporally quick signals in the PMTs.
The second T1 trigger is a time-over-threshold trigger, ToT-T1, requiring two PMTs to coincidentally measure above 0.2$\,I_{\rm VEM}^{\rm peak}$ in at least 13 FADC bins (\qty{325}{\nano\second}) over a 120 FADC bin (\qty{3}{\micro\second}) sliding window.
The ToT-T1 trigger reduces the trigger rate to $\approx$\qty{2}{\hertz} and is more sensitive to electromagnetic particles, as the Cherenkov light generated by these particles in the WCD is spread in time.
ToT triggers are automatically promoted to T2 triggers, whereas the Th-T1 trigger is only promoted to a Th-T2 if all three PMTs measure in coincidence a signal above 3.2$\,I_{\rm VEM}^{\rm peak}$.
The Th-T2 further reduces the trigger rate from \qty{100}{\hertz} to \qty{20}{hertz}.
Over the 20 years of Auger data taking, PMT failures have occurred in individual WCDs, with these PMTs referred to as ``masked'' PMTs.
In these conditions, the thresholds for Th and ToT triggers change to the values described in AUGER TRIGGER AND APERTURE, as a three PMT coincidence is no longer viable.

All station level T2 triggers are transferred to CDAS, where the T3 trigger is formed from a time and space coincidence of stations with a T2 trigger.
Once a T3 trigger is formed, a T3 request is sent to all stations participating in the T3 trigger and all nearby stations.
The data for each station participating in the T3 request is sent to CDAS where it is stored for further physics analyses.
The T4 trigger, the first physics trigger of the SD, requires stricter time and space coincidence of the stations than those used in the T3 trigger.
Stricter coincidence conditions ensure the events passing the T4 trigger are true physics events rather than random coincident triggering of nearby stations.
The T5 trigger, the final SD trigger, applies a fiducial quality cut to the true air-shower events.
A common fiducial cut to use is a 6T5 cut, also known as a 6T5 trigger, which requires the station with the highest signal in an event to be fully surrounded by six working stations at the time of the event.
This cut selects events fully contained within the SD array to ensure they have a reliable SD reconstruction and is a common cut to use for high-level SD analyses.


%used to select high quality 


%selects true physics events from the T3 triggered events 



%While stored at CDAS, two additional triggers are applied


%T1 and T2 triggers, respectively the first and second level triggers, are generated locally by the station electronics.
%Two types of T1 trigger, designed for sensitivity to different particle types, exist.



%The first two trigger levels, T1 and T2 respectively, are generated 


%data recorded with the local station \texttt{muon_buffer} used to calibrate the detector from atmospheric muons arriving at a rate of $\approx$\qty{3}{\kilo\hertz} at each WCD.


%, triggering once an individual PMT measures a peak signal above \qty{0.1}{VEM}.
%This signals are generated from atmospheric muons of low-energy air showers and trigger

%signals measured from atmospheric muons that form the charge calibration histograms of the individual PMTs.
%Atmospheric muons from low-energy air showers 
%A WCD equipped with an RDC hodoscope was used to experimentally confirm the conversion (REF 6 Calibration).





%Signals generated by the atmospheric muons form the charge calibration histograms of the individual PMTs used to determine the local detector conversion from electronics units to VEM.


%from the measured charge calibration histograms of the individual PMTs.
%Atmospheric muons from low-energy air showers are used to build the local charge histograms



%The signal measured by the PMTs is recorded in units of Vertical Equivalent Muons (VEM), where one VEM is represents the total signal deposited by a single muon traversing the WCD vertically and centrally.
%Conversion of the integrated FADC units (i.e. the electronics units) to VEM provides a common reference unit across the array which is independent of the individual detectors. 



%The PMTs measure both direct and reflected Cherenkov radiation.




\section{Fluorescence Detector}\label{sec:fd}

The Fluorescence Detector (FD) of the Pierre Auger Observatory consists of 27 fluorescence telescopes to observe the longitudinal development of cosmic-ray air showers above approximately \qty[parse-numbers=false]{10^{18}}{\electronvolt} in energy.
As an air shower develops, the ultra-relativistic charged secondary particles will excite nitrogen molecules in Earth's atmosphere, which, upon de-excitation, generate fluorescence light measurable by the fluorescence telescopes.
The telescopes are split between four locations surrounding the SD array: Los Leones, Los Morados, Loma Amarilla, and Coihueco.
Each location hosts six fluorescence telescopes, with the exception of Coihueco which hosts three additional fluorescence telescopes for observing air-showers at lower energies.
These three telescopes will be described later in the text.
The six telescopes at each location cover a combined \qty{180}{\degree} azimuthal field of view overlooking the SD array, with individual telescopes providing an azimuthal and elevation field of view of \qty{30}{\degree}$\,\times\,$\qty{30}{\degree}.
Operation of the telescopes is possible only on clear, moonless nights, without rain or strong winds, limiting the uptime of the FD to approximately 15\%.

The fluorescence telescopes follow a Schmidt optics design to minimize coma aberrations, arising from incoming light rays not being parallel to the axis of the mirror.
The basics of the telescope system consist of a spherical aperture with a UV filter and Schmidt corrector ring, a segmented mirror, and a camera of PMTs.
Fluorescence light enters the aperture with \qty{2.2}{\meter} diameter where it is filtered to UV wavelengths to reduce the amount of background light reaching the camera.
The filtered light is collected by a segmented mirror of nearly \qty{13}{\meter^{2}} in size and focused to the telescope's camera, composed of 440 total PMTs arranged in a 22$\,\times\,$20 grid.
The camera PMTs of each telescope connect to front-end electronics boards that handle the digitization of the PMT signals and the first and second level hardware triggers of the FD telescope. 
A schematic of a single fluorescence telescope is shown in FIGURE.

Three additional fluorescence telescopes, known as the High Elevation Auger Telescopes (HEAT), were installed next to the Coihueco FD site in 2009.
HEAT is similar to the other fluorescence telescopes, but the main difference is each telescope of HEAT can observe in either an upwards or downwards configuration.
In the upwards configuration, the telescopes of HEAT are pointed upwards by \qty{29}{\degree} in elevation, providing an elevation field of view of approximately \qtyrange{30}{58}{\degree}.
The downwards configuration has the same field of view in elevation as the other 24 fluorescence telescopes.
Elevating the telescope field of view allows measurements of the longitudinal development of lower energy air showers that reach their maximum development at higher altitudes, with HEAT lowering the fluorescence detection threshold to approximately \qty[parse-numbers=false]{10^{17}}{\electronvolt} in energy.
The HEAT field of view overlooks the SD-750 and SD-433 arrays, delivering high-quality hybrid detections of air showers at energies near and between the second knee and ankle of the cosmic ray energy spectrum.


%to provide an elevation field of view of approximately 

%HEAT extends the fluorescence measurements down  




%The light is focused from a segmented mirror of nearly \qty{13}{\meter^{2}} in size to the telescope camera.
%The camera

%The spherical aperture  

%A schematic of a single fluorescence telescope is shown in FIGURE.  

%The fluorescence telescopes follow a Schmidt optics design, used to observe wide field of view's while limiting aberrations. 


%The fluorescence light passes through a UV filter on the at each telescope location before reachn 


%Close to 90\% of the air-shower energy is 
%The total fluorescence light generated is proportional to the energy deposited by the secondary air-shower particles in the atmosphere, and therefore serves as a 







%observe the fluorescence light emitted by the de-excitation of nitrogen molecules in Earth's atmosphere, 
 

%monitor the atmosphere directly above the SD array
%The telescopes 
%When a cosmic-ray air shower 


%pointed inwards from the array edges to overlook the SD array.
%The telescopes monitor the atmosphere directly above the SD array 

%from the SD array edges to observ.
%to observe the atmosphere 

\section{Event Reconstruction}\label{sec:reconstruction}

The event reconstruction for the Pierre Auger Observatory occurs separately for the SD and FD, and is later combined for hybrid air-shower analyses.
All reconstruction occurs within Offline, the observatory's software package for analyzing air-shower data.
For the SD reconstruction, two main measurements are needed for each station that recorded event data: the start time of the signal in the station, to determine the geometry of the arriving cosmic-ray air shower, and the total signal in the station's PMTs, to determine an event-level observable related to the air-shower energy.
Determination of the time range of the event in the station and the conversion of the PMT signals from FADC units to VEM must be done before analyzing the station data for event-level analyses.
The time range is determined using the procedure described in REF2020JINST, and the conversion of FADC to VEM is performed using the SD calibration histograms described in~\cref{sec:sd_triggers}, with further information on the conversion provided in REF.

%Within a station, the signal baseline of all PMTs (both high and low-gain channels) must first be subtracted.
%A total baseline is determined by linearly interpolating between constant baseline segments to account for the affect of the signal on the baseline.
%Baseline segments are defined as at least \qty{1}{\micro\second} of consecutive FADC bins (40 bins) within a signal range of $\bar{b}\pm\sigma_{b}$, with $\bar{b}$ the average baseline in the segment and $\sigma_{b}$ nominally set to 2 FADC counts (see REF2020JINST for further details).
%After baseline subtraction, the PMT signals are converted from FADC units to VEM.   

Signal time windows are determined on a PMT-to-PMT basis.
Therefore, the nominal start and stop times for the station are taken from the station PMT with the longest signal time window.
The station's PMTs are averaged bin-by-bin, where the averaged PMT signal will be integrated over the nominal time window of the signal to determine the total signal measured in VEM within the SD station.
When the station is close to the air-shower axis, large signals are measured in the tank which can saturate either the high-gain, low-gain, or both channels of the PMT.
The saturation is dependent on both the distance to the air-shower axis and the energy of the arriving shower (REF 2020JINST).
In the case of no PMT saturation, the high-gain PMT channels are used to determine the total measured PMT signal, whereas in the case of high-gain saturation, the low-gain PMT channels are used instead.
In the case of both high-gain and low-gain saturation, the station is labeled as a saturated station.
Software within the Offline framework is used to recover the signals of such saturated stations for analysis (REF GAP2008-030, Darko ICRC2013).
However, identifying the signal window of the individual PMTs is always done with the high-gain PMT channel.

The air-shower geometry (point of impact on the ground and arrival direction) are reconstructed from an iterative procedure involving both the signal start times and total signals of the SD stations participating in the event.
A signal-weighted barycenter, determined from the center-of-mass of stations in the event, is used as the first estimate of the impact point on ground.
Assuming speed of light propagation of the particles, a plane wave fit to the start times of the three stations with the largest signal closest to the barycenter determines the first estimate of the arrival direction.
All stations with start times outside of the range \qtyrange{-2}{+1}{\micro\second} from the arrival time determined by the fit are labeled as accidental stations, stations that do not participate in the event yet were triggered, likely from accidental muons.
Stations participating in the event are fit using a chi-squared minimization process to a timing model of the wavefront arrival that accounts for wavefront curvature.
Four values are fit: the directional cosines, $u$ and $v$, of the unit vector describing the shower arrival direction (i.e. $\hat{a} = u\,\hat{x} + v\,\hat{y} + \sqrt{1 - u^{2} - v^{2}}\,\hat{z}$), the time the air-shower core intersects with the ground ($t_{c}$), and the radius of curvature of the shower front, where the radius of curvature is only used as a free parameter if at least 5 stations are participating in the event.
Otherwise, it is parameterized from events where the radius of curvature could be determined from the fit (REF2020JINST).

Once the shower timing and wavefront properties are determined, the shower size at an optimal distance to the air-shower axis, i.e. $S(r_{\rm opt})$, is determined from a fit to the lateral distribution function (LDF) of the station signals.
The LDF describes the measured station signals as a function of the station distance to the shower axis.
The SD reconstruction of the Auger Observatory uses a modified Nishimura-Kamata-Greisen (NKG) function for the LDF fit:

\begin{equation}
    S(r) = S(r_{\rm opt})\,\left( \frac{r}{r_{\rm opt}} \right)^{\beta}\,\left( \frac{r\,+\,r_{\rm s}}{r_{\rm opt}\,+\,r_{\rm s}} \right)^{\beta + \gamma}, \label{eq:modified_NKG_ldf}
\end{equation}

\noindent
with $r_{\rm s}$ a scale distance fixed to \qty{700}{\meter}.
$\beta$ and $\gamma$ are handled with data-driven parameterizations, dependent upon energy and zenith angle, unless their respective ``lever-arm'' criteria described in REF2020JINST are met.
If this criteria is met for $\beta$, it is treated as a free parameter in the LDF fit, and similarly for $\gamma$.

$S(r_{\rm opt})$ is the shower size estimated for the fit at the reference distance $r_{\rm opt}$.
Note that given the same energy, mass, and arrival direction of a primary particle, shower-to-shower fluctuations in the air-shower development lead to the air shower being measured at ground in different stages of the shower development.
The impact is an intrinsic variation in $S(r_{\rm opt})$.
Therefore, the reference distance was determined to minimize the variations in the shower-size estimator arising from shower-to-shower fluctuations and depends on the spacing of the SD array (REF24IN2020JINST).
ADDASFOOTNOTE For a general particle detector array, the reference distance also depends on the array structure, which for the Auger SD is an isometric triangular grid.
The optimal reference distances for the SD-1500, SD-750, and SD-433 arrays are \qtylist{1000;450;300}{\meter} respectively.
In each case, $S(r_{\rm opt})$ is a free parameter in the LDF fit.
The fit itself determines $S(r_{\rm opt})$ and the position of the air-shower core intersection with ground ($\vec{x}_{\rm c}$) by minimizing a log-likelihood function of the form:

\begin{equation}
    \ln\mathcal{L} = \sum_{i}\,\ln P\left( S(r_{\rm opt}),\vec{x}_{\rm c} | S_{i},\vec{x}_{i} \right), \label{eq:sd_ll}
\end{equation}

\noindent
with $S_{i}$ and $\vec{x}_{i}$ the measured signal and positions of the $i$-th station.
The probability model, $P$, depends on the signal measured by the station, with each model described in REF2020JINST.

For a given air-shower energy, less particles will reach the ground for inclined air showers than for vertical air showers due to attenuation effects as the air shower develops.
This directly affects the signal measured by the WCDs of Auger, introducing a zenith angle dependence on the reference distance signal, $S(r_{\rm opt})$.
To correct for this dependence, the constant intensity cut (CIC) method is used to derive an attenuation correction as a function of zenith angle, i.e. $f_{\rm att}(\theta)$ REFGAP2015-065.
The correction is applied following equation

\begin{equation}
    S_{\theta_{\rm ref}} = \frac{S(r_{\rm opt})}{f_{\rm att}(\theta)}, \label{eq:Sref_correction}
\end{equation}

\noindent
with $f_{\rm att}(\theta)$ a third-degree polynomial in $x = \cos^{2}(\theta) - \cos^{2}(\theta_{\rm ref})$.
The applied correction represents the average signal at reference distance $r_{\rm opt}$ measured if the incoming cosmic ray arrived at a zenith angle of $\theta_{\rm ref}$.
The reference angle changes depending upon the spacing of the SD array, with reference angles of \qtylist{38;35}{\degree} respectively used for the SD-1500 and SD-750 arrays REFSARXIV2008.06486, REFARXIV2109.13400.
The attenuation corrected signals are referred to as $S_{38}$ and $S_{35}$, and can be used in conjunction with hybrid FD measurements to derive an energy estimate from the SD.

For the FD reconstruction, the shower-detector plane, the plane connecting the triggered pixels of the telescope and the air-shower axis, is first determined using the timing information of the triggered fluorescence telescope pixels.
Uncertainties in the shower-detector plane reconstruction are validated with laser measurements from the Central Laser Facility (CLF) located within the SD-1500 array, as the distance and direction of the CLF laser is known with an accuracy better than the FD reconstruction resolution.
In the case of hybrid events where SD stations also participate in the event the shower-detector plane geometry is more precisely constrained than when no SD reconstruction is available.
For a majority of FD events, an SD reconstruction is available due to the approximate 100\% duty cycle of the SD.
For these hybrid events, resolutions of \qty{50}{\meter} for the air-shower core and \qty{0.6}{\degree} for the air-shower arrival direction are possible REFARXIV0907.4282.

The longitudinal development of the air shower, including the location of the depth of shower maximum ($X_{\rm max}$), can be reconstructed with the FD.
The time-dependent fluorescence light signals collected by the telescope pixels are converted to energy deposited by the shower in a portion of atmosphere, $\rm{d}E / \rm{d}X$, as a function of atmospheric slant depth, $X$, and fit to a Gaisser-Hillas function

\begin{equation}
    f_{\rm GH}(X) = \left( \frac{\rm{d}E}{\rm{d}X} \right)_{\rm max}\,\left( \frac{X - X_{0}}{X_{\rm max} - X_{0}} \right)^{(X_{\rm max} - X_{0}) / \lambda}\,\exp\left(\frac{X_{\rm max} - X}{\lambda}\right). \label{eq:gh}
\end{equation}

\noindent
$\left(\frac{\rm{d}E}{\rm{d}X}\right)_{\rm max}$ is the energy deposit in the atmosphere at shower maximum, i.e. $X_{\rm max}$. 
$X_{0}$ and $\lambda$ are shape parameters of the longitudinal profile, with $X_{0}$ related to the point of first interaction in the atmosphere and $\lambda$ representing an effective interaction length.
The shape parameters, along with $X_{\rm max}$ are parameters of the fit determined by the FD reconstruction.
An alternative Gaisser-Hillas function as a function of a translated slant depth, $X\prime = X - X_{\rm max}$, can be used in the reconstruction which removes the correlation between the shape parameters, instead relating them to physical properties in the shower profile as the shower width ($L$) and asymmetry ($R$) REFARXIV1811.04660. 
Integrating~\Cref{eq:gh} yields an estimate of the calorimetric energy of the air-shower representing roughly 85\% of the total air-shower energy.
The remaining 15\% of the total energy is referred to as the ``invisible'' energy, representing the energy in the shower carried by high-energy muons and neutrinos, and is not measurable with the fluorescence technique.
A data-driven analysis determines the invisible energy and is used to correct the measured calorimetric energy of the FD to obtain an estimate for the air-shower energy REFARXIV1901.08040.

For hybrid events, the attenuation corrected SD signals from~\Cref{eq:Sref_correction} ($S_{38}$ for the SD-1500 and $S_{35}$ for the SD-750) can be related to the FD reconstructed energy following

\begin{equation}
    E_{\rm FD} = A\,\left( \frac{S_{\theta_{\rm ref}}}{\rm VEM} \right)^{B}. \label{eq:energy_scale}
\end{equation}

\noindent
A log-likelihood fit is performed to determine the $A$ and $B$ fit parameters REFARXIV1503.09027 using a selection of high-quality hybrid events.
The resulting $A$ and $B$ fit parameters for the SD-1500 are presented in REFPoS(ICRC2019)450, and for the SD-750 are presented in REFARXIV2109.13400.
The data-derived relationship between the FD energy scale and the SD reference signal is used to set the SD energy scale, even when FD data are not available.

%The calorimetric energy is then corrected for the ``invisible'' energy of the shower in a data-driven

%from shower particles that do not deposit their energy in the atmosphere through the fluorescence mechanism, suc.

%The applied correction yields the signal at a reference angle $\theta_{\rm ref}$, which represents the 

%The reference angle depends upon the spacing of the SD array, as it is chosen to be the median zenith angle for the distribution of cosmic rays measured with the array.
%The SD-1500 measured cosmic rays 
%is the median zenith angle for the distribution of cosmic rays measured with the array, and therefore depends on the spacing of the SD array

%After the depth of shower maximum is reached for an air shower, its development in the atmosphere is dominated by attenuation effects

%The signal measured by the SD is sensitive to attenuation effects, as the SD stations measure less signal for inclined air showers than vertical air showers of the same energy.

%The reference distance signal, $S(r_{\rm opt})$ is sensitive to attenuation effects, must be corrected for zenith angle dependencies 



%The sampling of the LDF is sparse due to the detector spacing constraints of the SD, leading to imprecise constraints on the LDF shape on an event level.
%To offset this, the fit is performed to an average LDF, $f_{\rm LDF}(r)$:

%The nominal case is that of no PMT saturation, in which the high-gain PMT channels are used for the total signal measurements  
%In the case of high-gain saturation, the low-gain PMT channels are used to determine the total measured signal.
%In the case of both the high-gain, and low-gain 

%The PMTs can saturate if too much signal is measured within the station, typically oc 

%The PMT signals are then averaged within this time window over all working PMTs within the station


\section{Auger Engineering Radio Array}\label{sec:aera}

The Auger Engineering Radio Array (AERA) consists of 153 radio antennas deployed over a \qty{17}{\kilo\meter^{2}} area to observe the radio emission of air showers within a frequency band of \qtyrange{30}{80}{\mega\hertz}.
AERA is designed to measure the radio emission of air showers above approximately \qty[parse-numbers=false]{10^{17}}{\electronvolt} in energy and is co-located with the SD-750 array and within the field of view of the Coihueco FD site, providing coincident radio detections of air showers with the SD and FD.
AERA stations are autonomously powered with solar panels, providing an approximate 100\% duty cycle of the antennas.
Further, each station is equipped with a GPS system for timing information, a \qty{5}{\giga\hertz} wireless radio system for data transfer and communications, and fencing to prevent interference from wildlife.
Two types of antenna were used for AERA, the log-periodic dipole antenna (LPDA) and the butterfly antenna.
Both have two orthogonal polarizations, one aligned in the North-South direction (aligned with the local magnetic field) and the other in the East-West direction.
The signal processing chain for the LPDA and butterfly antennas is similar (described in REFAERA), yet differences arise within the responses of the antennas to air-shower radio emission.
LPDA antennas are more sensitive than the butterfly antennas, and therefore have a lower signal threshold, while the butterfly antennas have a signal dependence on local ground conditions, as this antenna's total measured signal is a combination of signals reflected from the ground that interfere constructively with the direct shower radio emission.
Both antennas are more sensitive to ``vertical'' air showers, i.e. air showers with arrival directions of \qtyrange{0}{55}{\degree} in zenith angle.

%The response of the LPDA and butterfly antennas to air-shower radio emission is mostly similar, with the exception that LPDA antennas are more sensitive to the butterfly antennas and therefore 
%The two antenna types both measure the radio emission from air showers within a frequency band of \qtyrange{30}{80}{\mega\hertz}.

AERA deployment started in 2011 and was performed over three phases.
AERA phase I consists of 24 LPDA antennas on a triangular grid with inter-antenna spacing of \qty{144}{\meter}.
Phase II deployment started in 2013, adding an additional 100 butterfly antennas: 60 deployed on a \qty{250}{\meter} grid and the remaining 40 deployed on a \qty{375}{\meter} grid.
In 2015, AERA phase III consisting of 29 butterfly antennas was deployed on a \qty{750}{\meter} grid, resulting in the completion of AERA.

Two trigger types exist for the AERA stations: an external trigger, supplied by local detector triggers from CDAS (mostly triggers from the SD), and a radio self trigger, that triggers directly on the radio signal itself.
The station electronics differ depending on the type of trigger algorithm used by the antenna.
Externally-triggered stations continually measure and digitize radio signals with a 12-bit digitizer at a \qty{180}{\mega\hertz} sampling rate.
The signals are then stored on a ring buffer of \qty{7.4}{\second} where they wait for the arrival of an external trigger from CDAS, taking about \qty{6}{\second} to arrive at the AERA stations in the case of SD generated triggers (Ricardo private communication). % Include??
The GPS timestamp of the trigger is used to readout the ring buffer data where it is transferred to the AERA data acquisition system at the Coihueco FD site, either directly from the antenna itself or from a Ubiquiti Bullet link at the CRS.
The self-triggered stations measure and digitize radio signals with a 14-bit digitizer at a \qty{200}{\mega\hertz} sampling rate.
The digitized signals are passed to an electronics board for threshold and signal determinations over the omnipresent radio background, where radio signals are sent to the CRS if the self-trigger signal conditions are met.
Since January 2023, all self-triggered AERA stations have been decommissioned, 10 of which have been repurposed for GRAND@Auger (REFGRAND@AUGER).

A major science goal of AERA was to serve as a testbed for deploying a radio array in the \qtyrange{30}{80}{\mega\hertz} band over the full surface array of the Auger observatory as part of it's next generation upgrade.
However, high-quality radio data and important analyses with AERA strengthened the cosmic-ray science potential of the future radio array.
Science highlights include an $X_{\rm max}$ reconstruction with AERA (PRDPRL BJARNI), deriving an absolute energy scale from radio measurements (ICRC2025TIM), and a Galactic calibration of the AERA antennas proving their long-term stability (CITEGALCALPAPER).
A recent analysis combines AERA radio measurements, which probe the electromagnetic air-shower component, with SD reconstructions of inclined ($\theta_{\rm zen}\,>\,$\qty{60}{\degree}) to provide a simultaneous muon and radio measurement on a per-shower basis and serves as a proof of principle for the radio component of the Auger observatory's next generation upgrade (see~\Cref{sec:rd} for further details).
The aim of this thesis is to study simultaneous muon and radio measurements in the context of vertical ($\theta_{\rm zen}\,<\,$\qty{55}{\degree}) air showers measured by the Auger detector systems.

%However, AERA not only improved the understanding of the radio detection technique for air showers in this frequency range in Argentina, but also important science results with AERA strengthened the cosmic-ray science potential of the future radio array.

%AERA was used to prove in principle the radio detection technique at the Auger observatory in the frequency band of \qtyrange{30}{80}{\mega\hertz} and to serve as a testbed for a radio array to be deployed over the full surface array of the Pierre Auger Observatory as part is the observatory's next generation upgrade.

%prepare for the next generation upgrade of the Pierre Auger Observatory, AugerPrime, 

\section{AugerPrime: The Upgrade of the Pierre Auger Observatory}\label{sec:augerprime}

AugerPrime is the next generation upgrade to the Pierre Auger Observatory in the quest to determine the astrophysical origin of ultra-high-energy cosmic rays.
The main science objective of AugerPrime is to improve the observatory's composition sensitivity, to be achieved with more stringent separation of the electromagnetic and muonic air-shower components.
Independent measurements of the muonic component are crucial, as the muon puzzle highlights the incompleteness of our interaction models in extrapolated energy ranges (see~\Cref{sec:models}), and directly serve another science goal of AugerPrime: studying and constraining hadronic interaction models. 
To this end, the observatory has upgraded its current detectors, in addition to adding new detector systems over the full array.
Deployment of all AugerPrime components finished in March 2026.

Each WCD of the SD was outfitted with a surface scintillator detector (see~\Cref{sec:ssd}) and a radio antenna (see~\Cref{sec:rd}), both deployed on top of the WCDs and directly integrated into the SD detector system and electronics.
A small PMT (SPMT), a Hamamatsu R8619 PMT of 1-inch diameter, was installed inside each WCD to increase the dynamic range of the SD, allowing SD measurements without PMT saturation to within \qty{250}{\meter} of the reconstructed air-shower core REFICRC2023343.
In addition, the electronics of the SD were updated with an Upgraded Unified Board (UUB).
Compared to the UB, the UUB increases the dynamic range from 10 to 12 bits and increases the sampling rate from \qty{40}{\hertz} to \qty{120}{\hertz}, respectively resulting in improved signal amplitude and timing accuracy of the PMT signals REFARXIV1604.03637.
Other electronics updates include improvements to the processing power, local stations monitoring system, and the local trigger and calibration capabilities.
See REFJINST202318P10016 for further information on the SD electronics of AugerPrime.
Deployment of the UUB electronics was completed in June 2023.
AugerPrime has also improved the duty cycle of the FD up to 29\% by lowering the PMT gain.
Lowering the gain allows observation nights with moonlight, i.e. nights with a higher night sky background, without risk of deteriorating the FD PMTs from high anode currents.
The new detector systems added as part of the AugerPrime upgrade will be discussed in the following subsections.

%closer to the shower core without PMT saturation REFICRC2023343.
%The electronics used for the SD array will also be up

%However, solely increasing statistics of air-shower observations 

%The science objective of AugerPrime is to determine with precision the mass composition of ultra-high-energy cosmic rays, with focus on improving the observatory's composition sensitivity   


\subsection{Underground Muon Detector}\label{sec:amiga}

\subsection{Surface Scintillator Detector}\label{sec:ssd}

\subsection{Radio Detector}\label{sec:rd}

